Interactive Data Exploration with Smart Drill-Down

### 2017-03-10 TableSummarization/journalVersion/smartDrillDown.tex
Addressed comments, changed title, change image formats

\title{Interactive Data Exploration with Smart Drill-Down (Extended Version)}

@@

\begin{figure}
\centering
\includegraphics[width=80mm,frame]{graphs/screenshots6col/empty_expansion.jpg}
\caption{Summary after clicking on the empty rule \label{fig:uiexample1}}
\vspace{-10pt}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=80mm,frame]{graphs/screenshots6col/education_expansion.jpg}
\caption{Star expansion on `Education' Column \label{fig:uiexamplestar}}
\vspace{-10pt}
\end{figure}

@@

\techreporttext{
\begin{figure}
%\vspace{-10pt}
\centering
\includegraphics[width=80mm,frame]{graphs/screenshots6col/rule_expansion.jpg}
\caption{A rule expansion \label{fig:uiexamplerule}}
\vspace{-10pt}
\end{figure}
}

\begin{figure}
\centering
\includegraphics[width=80mm,frame]{graphs/screenshots6col/age_drilldown.jpg}
\caption{A regular drill down on Age\label{fig:drilldownexample}}
\vspace{-10pt}
\end{figure}

@@

\techreporttext{
\begin{figure}
\centering
\includegraphics[width=80mm,frame]{graphs/screenshots6col/bits_scoring.jpg}
\caption{Bits scoring\label{fig:weigtingbitwise}}
\end{figure}

\begin{figure}
\vspace{-10pt}
\centering
\includegraphics[width=80mm,frame]{graphs/screenshots6col/min2_scoring.jpg}
\caption{Size minus one weighting \label{fig:weightingsizeminusone}}
\vspace{-10pt}
\end{figure}
}

@@

\papertext{
\vspace{-5pt}
\section{Extensions}\label{sec:extensions}
\papertext{
\subsection{Setting parameters $W$, $k$, $m_w$}\label{sec:setting-paramters}
Our system allows the user to tune the smart drill-down by adjusting a number of parameters. Having a lot of tunable parameters can increase the difficulty of using a system by increasing the decision-making burden on the user. To counteract this, we now provide ways to guide the user while selecting appropriate parameter values.

@@

\noindent
\textbf{Generalizing our weight functions:} We now analyze a parametric family of weighting functions that generalizes our functions from Section~\ref{sec:weighting} and provides intuition for them. Let $W(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^k$, where the $w_c$'s and $k$ are parameters of the weighting function, and $o_{r,c}$ ($o$ stands for occurrence) is $1$ if $r(c) \neq \star$ and $0$ otherwise. The Size and Bits weighting functions can be seen as special cases of the above function with $k=1$ and with $w_c$ equal to $1$ for Size and $\text{log}(|c|)$ for Bits.

We want to estimate the Score contribution of a rule according to this weight function, which requires estimating the count of the rule. Suppose that for each column $c$, the most frequent value occurs a $f_c$ fraction of the time. If the probability of a value occurring in a column of a tuple is independent of other values in the tuple, the $Count$ of rule $r$ is approximately $|T|\Pi_{c \in C}f_c^{o_{r,c}}$. Thus the score of a rule set containing $r$ alone can be approximated by:
$$S(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^k|T|\Pi_{c \in C}f_c^{o_{r,c}}$$

Now consider the rule that has the highest weight times count according to the above weighting function. We can approximate this rule by relaxing $o_{r,c}$ to a real number; that is by maximizing the estimated $S(r)$ subject to $0 \leq o_{r,c} \leq 1$. According to the KKT conditions~\cite{Boyd:2004:CO:993483}, at the optimum $Weight \times Count$, the partial derivative of $W(r)$ with respect to $o_{r,c}$ must be $\leq 0$ for each $c$ where $o_{r,c} = 0$, equal to $0$ if $0 < o_{r,c} < 1$ and $\geq 0$ otherwise. For each $c$, we have:
$$\frac{\partial S(r)}{\partial o_{r,c}} = S(r)(\frac{kw_c}{\sum_{c \in C} o_{r,c}w_c} + \ln f_c)$$

Thus, if $x = \frac{k}{\sum_{c \in C} o_{r,c}w_c}$, then the maximum scoring rule should instantiate columns where $\frac{\ln f_c}{w_c} > x$ and some of the ones with $\frac{\ln f_c}{w_c} = x$. Thus, for Size weighting, the highest score rule should select columns with high values of $\ln f_c$, i.e., the highest frequency values. In contrast, Bits weighting will select columns with highest values of $\log_{|c|}(f_c)$, i.e., values that are frequent relative to the number of distinct values in the column.

@@

We can also estimate the size of the optimal rule under the above weighting function. Specifically, the weighted fraction of columns with non-$\star$ values is given by $\frac{\sum_{c\in C} o_{r,c}w_c}{\sum_{c \in C} w_c}$ which equals $-k/(\sum_{c \in C} \ln f_c)$. Thus, if we wanted our highest score rule to have $s$ fraction of the columns get instantiated, we can achieve that by setting $k = -s\sum_{c \in C} \ln f_c$.

The above reasoning also lets us estimate the $m_w$ parameter we need to use. Specifically, the highest score rule above has weight $(\frac{-\ln f_c}{kw_c})^k$.

>>>>

\title{Interactive Data Exploration with Smart Drill-Down}

@@

\begin{figure}
\centering
\includegraphics[width=80mm,frame]{graphs/screenshots6col/empty_expansion.png}
\caption{Summary after clicking on the empty rule \label{fig:uiexample1}}
\vspace{-10pt}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=80mm,frame]{graphs/screenshots6col/education_expansion.png}
\caption{Star expansion on `Education' Column \label{fig:uiexamplestar}}
\vspace{-10pt}
\end{figure}

@@

\techreporttext{
\begin{figure}
%\vspace{-10pt}
\centering
\includegraphics[width=80mm,frame]{graphs/screenshots6col/rule_expansion.png}
\caption{A rule expansion \label{fig:uiexamplerule}}
\vspace{-10pt}
\end{figure}
}

\begin{figure}
\centering
\includegraphics[width=80mm,frame]{graphs/screenshots6col/age_drilldown.png}
\caption{A regular drill down on Age\label{fig:drilldownexample}}
\vspace{-10pt}
\end{figure}

@@

\techreporttext{
\begin{figure}
\centering
\includegraphics[width=80mm,frame]{graphs/screenshots6col/bits_scoring.png}
\caption{Bits scoring\label{fig:weigtingbitwise}}
\end{figure}

\begin{figure}
\vspace{-10pt}
\centering
\includegraphics[width=80mm,frame]{graphs/screenshots6col/min2_scoring.png}
\caption{Size minus one weighting \label{fig:weightingsizeminusone}}
\vspace{-10pt}
\end{figure}
}

@@

\papertext{
\vspace{-5pt}
\section{Extensions}\label{sec:extensions}
\papertext{
\subsection{Setting parameters $k$, $m_w$, $W$}\label{sec:setting-paramters}
Our system allows the user to tune the smart drill-down by adjusting a number of parameters. Having a lot of tunable parameters can increase the difficulty of using a system by increasing the decision-making burden on the user. To counteract this, we now provide ways to guide the user while selecting appropriate parameter values.

@@

\noindent
\textbf{Generalizing our weight functions:} We now analyze a parametric family of weighting functions that generalizes our functions from Section~\ref{sec:weighting} and provides intuition for them. Let $W(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^{\mathsf{p}}$, where the $w_c$'s and $\mathsf{p}$ are parameters of the weighting function, and $o_{r,c}$ ($o$ stands for occurrence) is $1$ if $r(c) \neq \star$ and $0$ otherwise. The Size and Bits weighting functions can be seen as special cases of the above function with $\mathsf{p}=1$ and with $w_c$ equal to $1$ for Size and $\text{log}(|c|)$ for Bits.

We want to estimate the Score contribution of a rule according to this weight function, which requires estimating the count of the rule. Suppose that for each column $c$, the most frequent value occurs a $f_c$ fraction of the time. If the probability of a value occurring in a column of a tuple is independent of other values in the tuple, the $Count$ of rule $r$ is approximately $|T|\Pi_{c \in C}f_c^{o_{r,c}}$. Thus the score of a rule set containing $r$ alone can be approximated by:
$$S(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^{\mathsf{p}}|T|\Pi_{c \in C}f_c^{o_{r,c}}$$

Now consider the rule that has the highest weight times count according to the above weighting function. We can approximate this rule by relaxing $o_{r,c}$ to a real number; that is by maximizing the estimated $S(r)$ subject to $0 \leq o_{r,c} \leq 1$. According to the KKT conditions~\cite{Boyd:2004:CO:993483}, at the optimum $Weight \times Count$, the partial derivative of $W(r)$ with respect to $o_{r,c}$ must be $\leq 0$ for each $c$ where $o_{r,c} = 0$, equal to $0$ if $0 < o_{r,c} < 1$ and $\geq 0$ otherwise. For each $c$, we have:
$$\frac{\partial S(r)}{\partial o_{r,c}} = S(r)(\frac{\mathsf{p}w_c}{\sum_{c \in C} o_{r,c}w_c} + \ln f_c)$$

Thus, if $x = \frac{\mathsf{p}}{\sum_{c \in C} o_{r,c}w_c}$, then the maximum scoring rule should instantiate columns where $\frac{\ln f_c}{w_c} > x$ and some of the ones with $\frac{\ln f_c}{w_c} = x$. Thus, for Size weighting, the highest score rule should select columns with high values of $\ln f_c$, i.e., the highest frequency values. In contrast, Bits weighting will select columns with highest values of $\log_{|c|}(f_c)$, i.e., values that are frequent relative to the number of distinct values in the column.

@@

We can also estimate the size of the optimal rule under the above weighting function. Specifically, the weighted fraction of columns with non-$\star$ values is given by $\frac{\sum_{c\in C} o_{r,c}w_c}{\sum_{c \in C} w_c}$ which equals $-\mathsf{p}/(\sum_{c \in C} \ln f_c)$. Thus, if we wanted our highest score rule to have $s$ fraction of the columns get instantiated, we can achieve that by setting $\mathsf{p} = -s\sum_{c \in C} \ln f_c$.

The above reasoning also lets us estimate the $m_w$ parameter we need to use. Specifically, the highest score rule above has weight $(\frac{-\ln f_c}{\mathsf{p}w_c})^{\mathsf{p}}$.
